---
title: "p8105_hw2_xy2397"
author: "Xue Yang"
date: "9/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


# Problem 1

**Read and clean the data**

```{r}
NYC_transit_data = 
  read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:route11, entry, vending, entrance_type, ada, ada_notes) %>%
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

**Short paragraph about the dataset**

This dataset "NYC_Transit_Subway_Entrance_And_Exit_Data" (the original dataset before cleaning) contains variables: Division, Line, Station Name, Station Latitude, Station Longitude, Station Location (Station Latitude, Station Longitude), Route from 1 to 11, Entrance Type (stair, door, elevator, escalator or easement), Entry (yes or no), Entrance Latitude, Entrance Longitude, Entrance Location (Entrance Latitude, Entrance Longitude), Exit Only (whether or not), Vending (yes or no), Staffing (full, none or part), Staff Hours, ADA (ADA compliant or not), ADA Notes, Free Crossover (true or false), North South Street, East West Street, Corner(NW,NE, SW or SE).

I clean the data and rename the data "NYC_transit_data" from the following step:

* Import data: I import the data using relative path.

* Clean variable name: using clean_names() function from janitor package to clean up variable names and convert them to lower snake case after importing data

* Select variables: only retain the following variables: line, station name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance.

* Covert variable: convert the entry variable from character (YES vs NO) to a logical variable (TRUE vs FALSE).

Then the dimension of dataset after cleaning "NYC_transit_data" is `r nrow(NYC_transit_data)` rows x `r (ncol(NYC_transit_data))` columns.

These data is not tidy, since the data provided includes columns "route1" to "route11", which list all routes served by a station entrance/exit, which is not a "tidy" format.

**Answer the questions using the data**

•	How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St A/B/C/D; 125st 1; 125st 4/5); the distinct function may be useful here.

```{r}
distinct_station = 
  distinct(NYC_transit_data, line, station_name, .keep_all = TRUE)

nrow(distinct_station)
```


•	How many stations are ADA compliant?

```{r}
ADA_compliant =  filter(distinct_station, ada == "TRUE")

nrow(ADA_compliant)
```


•	What proportion of station entrances / exits without vending allow entrance?

```{r}
no_vending = filter(distinct_station, vending == "NO")

nrow(no_vending)/nrow(distinct_station)
```

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r}
NYC_transit_data %>% 
  gather(key = "route_number", value = "route_name", route1:route11) %>% 
  distinct(station_name, .keep_all = TRUE) %>% 
  filter(route_name == "A") %>% 
  nrow()
```

```{r}
NYC_transit_data %>% 
  gather(key = "route_number", value = "route_name", route1:route11) %>% 
  distinct(station_name, .keep_all = TRUE) %>% 
  filter(route_name == "A") %>% 
  filter(ada == "TRUE") %>% 
  nrow()
```


